---
title: "DecisionTree"
author: "leonardo_2368386"
date: "2024-03-23"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Decision Tree ML method

## Load libraries
```{r}
# install the rpart package (if needed) to create the decision tree model
if(require(rpart) == FALSE){
  install.packages('rpart')
  library(rpart)
}
# install the Metrics package (if needed) to evaluate the model's performance
if(require(Metrics) == FALSE){
  install.packages("Metrics")
  library(Metrics)
}
# install caret (if needed) to normalize the data
if(require(caret) == FALSE){
  install.packages("caret")
  library(caret)
}

```

## 1.Load datasets

```{r}

movies_dataset <- read.csv("movies_final.csv", stringsAsFactors = T)

#70/30 split
training_movies <- read.csv("training_movies.csv", stringsAsFactors = T)

testing_movies <- read.csv("testing_movies.csv", stringsAsFactors = T)

```

## 2.Data preparation

```{r}

# Removal of the title column
training_movies <- training_movies[,-1]
testing_movies <- testing_movies[,-1]

summary(movies_dataset)
str(movies_dataset)

summary(training_movies)
str(training_movies)

summary(testing_movies)
str(testing_movies)

print("number of rows in the training set")
nrow(training_movies)

print("number of rows in the testing set")
nrow(testing_movies)

```

## 3.Decision tree model training

```{r}

# Create decision tree model using regression 
#input variables:
# -Duration
# -Budget
# -Reviews_by_Users
# -IMDB_Score
#output variables:
# -Gross_Earnings
  
DT_model <- rpart(Gross_Earnings ~ Duration + 
             Budget + Reviews_by_Users + IMDB_Score,  
             method = "anova", data = training_movies)

```

## 4.Model view (Plots)

```{r}

plot(DT_model, uniform = TRUE, 
          main = "Gross_Earnings Decision  
                 Tree using Regression") 
text(DT_model, use.n = TRUE, cex = .7)

print(DT_model)

```

## 5.Decision Tree prediction

```{r}

#prediction of the testing set values
cat("Predicted value:\n") 
DT_pred <- predict(DT_model, testing_movies, method = "anova") 

```

## 6.Data Normalization

```{r}

# Testing set true "Gross_Earnings" values
actual_DT_values <- testing_movies$Gross_Earnings

#normalize the data using Min-Max Scaling to a scale of 0 to 1

#actual values normalization

actual_prep <- preProcess(as.data.frame(actual_DT_values), method=c("range"))

norm_actual_DT_values <- predict(actual_prep, as.data.frame(actual_DT_values))

#predicted values normalization

DT_pred_prep <- preProcess(as.data.frame(DT_pred), method=c("range"))

norm_DT_pred <- predict(DT_pred_prep, as.data.frame(DT_pred))

```

## 7.Decision Tree performance evaluation

```{r}

#To evaluate the model's performance we are going to use metrics such as root mean squared error (RMSE) and the R-squared value

# Calculate RMSE and R-squared

#plot actual values vs predicted values
plot(norm_actual_DT_values$actual_DT_values, norm_DT_pred$DT_pred, 
     xlab = "Actual Values", 
     ylab = "Predicted Values", 
     main = "Actual vs. Predicted Values")

abline(0, 1, col = "red")

legend("topleft", legend = "Ideal Fit", col = "red", lty = 1, cex = 0.8)

# Calculate RMSE
RMSE_value <- rmse(norm_DT_pred$DT_pred, norm_actual_DT_values$actual_DT_values)#rmse(DT_pred, actual_DT_values)

# calculate r-squared
rsq <- function(x, y) summary(lm(y~x))$r.squared
r2 <- rsq(norm_DT_pred$DT_pred, norm_actual_DT_values$actual_DT_values)#rsq(DT_pred, actual_DT_values)

# Print the metrics
cat("RMSE:", RMSE_value, "\n")
cat("R2:", r2, "\n")

```

  By looking at the results, we can say that the r-squared value obtained shows that approximately 53.15% of the variance in the dependent variable "Gross_Earnings" can be explained by the other independent variables used in the model, and the root mean squared error (RMSE) showed a result value of 0.17. Therefore, we obtained considerably good results with these metrics.

#based on the following sources:
-https://www.datacamp.com/tutorial/decision-trees-R
-https://www.geeksforgeeks.org/decision-tree-for-regression-in-r-programming/
-https://www.digitalocean.com/community/tutorials/normalize-data-in-r
