---
title: "CS5812-Dua-NN"
output: html_document
date: "2024-03-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install and load the neuralnet package from CRAN
if(require(neuralnet) == FALSE){
  install.packages('neuralnet')
  
  
}
# install the Metrics package (if needed) to evaluate the model's performance
if(require(Metrics) == FALSE){
  install.packages("Metrics")
  library(Metrics)
  
}

# install caret (if needed) to normalize the data

  #install.packages("caret")
  library(caret)

```

#### 2. Data preparation
```{r}

# read the data from the movies.csv file
movies_df <- read.csv("movies_final.csv")

### 1.2 load the test and train movies data
movies_test <- read.csv("testing_movies.csv")
movies_train <- read.csv("training_movies.csv")

# inspect the data
str(movies_df)
training_movies<-subset(movies_train, select = -Title)
test_movies<-subset(movies_test, select = -Title)

```
#### 3. Data Normalization
```{r}
movies_test_prep <- preProcess(as.data.frame(test_movies), method=c("range"))
test_movies_minmax <- predict(movies_test_prep, as.data.frame(test_movies))

movies_train_prep <- preProcess(as.data.frame(training_movies), method=c("range"))
training_movies_minmax <- predict(movies_train_prep, as.data.frame(training_movies))
```
#### 4. Neural network training{#Neural_network_training}
```{r}
# define a formula for predicting gross_earnings
grossearning_formula = Gross_Earnings ~ Duration+Budget+Reviews_by_Users+IMDB_Score

# train a neural network with 1 hidden node
gs_nn_1 <- neuralnet(grossearning_formula , data = training_movies_minmax)

# train a neural network with 4 nodes on one hidden layer
gs_nn_4 <- neuralnet(grossearning_formula , hidden = 4, data = training_movies_minmax)

# train a neural network with 4 nodes on each of two hidden layers
gs_nn_44 <- neuralnet(grossearning_formula , hidden = c(4,4), data = training_movies_minmax)

# plot the three neural networks and compare their structure
plot(gs_nn_1)
plot(gs_nn_4)
plot(gs_nn_44)
```

#### 5. Neural network prediction{#Neural_network_prediction}
```{r}
# compute the prediction for each neural network
pred_gs_nn_1 <- compute(gs_nn_1, test_movies_minmax[,-5])
pred_gs_nn_4 <- compute(gs_nn_4, test_movies_minmax[,-5])
pred_gs_nn_44 <- compute(gs_nn_44, test_movies_minmax[,-5])

# create a table with actual values and the three predictions
gs_results <- data.frame(
  actual = test_movies_minmax$Gross_Earnings,
  nn_1 = pred_gs_nn_1$net.result,
  nn_4 = pred_gs_nn_4$net.result,
  nn_44 = pred_gs_nn_44$net.result
)

# calculate the correlation between actual and predicted values to identify the best predictor
cor(gs_results[,'actual'], gs_results[,c("nn_1","nn_4", "nn_44")])

# plot actual vs predicted values for the worst (blue) and best predictor (orange)
#   note: points is used to add points on a graph
plot(
  gs_results$actual,
  gs_results$nn_44,
  col = 'blue',
  xlab = 'actual result',
  ylab = 'predicted result',
  xlim = c(0,1),
  ylim = c(0,1)
)
points(
  gs_results$actual,
  gs_results$nn_4,
  col = 'orange'
)
abline(a = 0, b = 1, col = 'red', lty = 'dashed')
legend(
  'topleft',
  c('nn_4', 'nn_44'),
  pch = 1,
  col = c('orange', 'blue'),
  bty = 'n'
)
```

#### 6. Performance evaluation

```{r}

# Calculate RMSE
RMSE_value_1 <- rmse( gs_results$actual,
  gs_results$nn_1)

# calculate r-squared
rsq <- function(x, y) summary(lm(y~x))$r.squared
r2_1 <- rsq( gs_results$actual,
  gs_results$nn_1)

# Print the metrics for 1 hidden layer
cat("RMSE:", RMSE_value_1, "\n")
cat("R2:", r2_1, "\n")
# Calculate RMSE
RMSE_value_4 <- rmse( gs_results$actual,
  gs_results$nn_4)

# calculate r-squared
r2_4 <- rsq( gs_results$actual,
  gs_results$nn_4)

# Print the metrics for 4 hidden layers
cat("RMSE nn_4:", RMSE_value_4, "\n")
cat("R2 nn_4:", r2_4, "\n")
# Calculate RMSE
RMSE_value_44 <- rmse( gs_results$actual,
  gs_results$nn_44)

# calculate r-squared 
r2_44 <- rsq( gs_results$actual,
  gs_results$nn_44)

# Print the metrics for two 4 hidden layers
cat("RMSE nn_44:", RMSE_value_44, "\n")
cat("R2 nn_44:", r2_44, "\n")


```

I trained three neural networks with different hyper-parameters. As I had 4 independent variables, I choose four neuron per hidden layer in order to train my model. The results that is r-squared value obtained by the performance evaluation of each trained model are as follows:
1. One Hidden node- 61%
2. One hidden layer with 4 nodes- 63%
3. Two hidden layers with 4 nodes each- 55%

Hence, we can conclude the neural network with one hidden layer and four neurons was the best among the three for predicting gross earning with accuracy 63%. 

